---
title: "Preprocessing Airbnb Melbourne Dataset"
author: "Yoga Pratama & Benjamin Huang"
output:
  html_document:
    df_print: paged
  html_notebook: default
---

## Required Packages 

```{r message = FALSE, warning = FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(rvest)
library(stringr)
library(lubridate)
library(forecast)
```


## Executive Summary 

For this project, open source data on Airbnb, along with data from a Wikipedia table on Local Government Areas (LGA) were used for the purpose of data pre-processing. The following steps were taken:

* Airbnb data was loaded and subset with the desired variables, while LGA data was scraped via Wikipedia and also subset with the desired variables. These two datasets were then relationally joined to form a new dataset.
* Variables in the new dataset were renamed to better reflect their meanings, and were then converted to more appropriate variable types.
* The dataset was briefly examined to check if it was in a ‘tidy’ format, in which it was hence no further action was taken regarding format.
* New variables were then created from the existing variables and added to the dataset.
* The entire dataset was scanned for NA values and other special values (infinite numbers and ‘Not a Number’ values), which were then removed according to appropriateness. 
* Rules were also defined and applied to check for any inconsistencies. For certain rule violations, observations were removed, while for other rule violations, observations were imputed. 
* The appropriate variables were scanned for outliers via Tukey’s method based on the distributions of the variables, which were then imputed due to the number of outliers that were found. Other methods were also briefly discussed.
* For the last step, one variable was chosen to be transformed and normalised in order to apply a z-score outlier detection method.

## Data 

The main dataset used for this project is the "Melbourne Airbnb Open Data" as hosted on Kaggle by Tyler Xie (https://www.kaggle.com/tylerx/melbourne-airbnb-open-data). The data was gathered from the Inside Airbnb database and compiled by Tyler Xie into a CSV file. There are 96 variables for the Airbnb dataset, however only 12 variables were chosen for the purpose of this project:

* id = Identification number for Airbnb listing.
* host_since = The date when the host of the lisitng signed up to Airbnb as a host.
* street = Street address.
* neighbourhood_cleansed = Local government area.
* property_type = The type of building (eg: House, Apartment, etc.)
* accommodates = The number of guests the accommodation can accommodate.
* bathrooms = The number of bathrooms.
* bedrooms = The number of bedrooms.
* price = Price per night.
* guests_included = The number of guests included in the price per night.
* number_of_reviews = The number of reviews the accommodation has.
* instant_bookable = Whether the accommodation instantly bookable.

Below, the Airbnb dataset is imported and subset according to the varibles chosen above:
```{r}

airbnb_orig <- read.csv("listings_dec18.csv", stringsAsFactors = F)
airbnb <- airbnb_orig[,c("id","host_since","street","neighbourhood_cleansed","property_type","accommodates",
                         "bathrooms","bedrooms","price","guests_included","number_of_reviews", "instant_bookable")]
head(airbnb)
```
\
The second dataset is scraped from a table off the Wikipedia article "Local Government Areas of Victoria" (https://en.wikipedia.org/wiki/Local_government_areas_of_Victoria#Greater_Melbourne). Its variables contain information on land size and population for each LGA. Only 3 variables were chosen out of the 11 variables, which were then renamed:

* LGA = Local government area.
* LGA_Area_km2 = The size of the LGA in kilometres squared.
* LGA_Density = The population density per kilometre squared.

Below, the Wikipedia table is scraped and imported using the 'rvest' package, and then subsetted with the chosen variables above. The variables are then renamed, while the LGA observations are edited to remove the strings "City of" or "Shire of" in front of the LGA names. This is so that the LGA variable will match with the 'neighbourhood_cleansed' variable in the Airbnb dataset so that the two datasets may be joined in the next step:
```{r}
LGA_wiki <- read_html("https://en.wikipedia.org/wiki/Local_government_areas_of_Victoria#Greater_Melbourne")
length(html_nodes(LGA_wiki, "table"))
LGA <- html_table(html_nodes(LGA_wiki, "table") [[1]], fill=T)
head(LGA)
LGA <- LGA[2:32,c(1,5,7)]
colnames(LGA) <- c("LGA","LGA_Area_km2","LGA_Density")
LGA <- LGA %>% 
  mutate(LGA = ifelse(str_detect(LGA, "Shire of"), str_replace(LGA, pattern="Shire of ", replacement=""), 
               ifelse(str_detect(LGA, "City of"), str_replace(LGA, pattern="City of ", replacement=""), "Error")))
head(LGA)
```
\
The two datasets are joined below, where Airbnb's 'neighbourhood_cleansed' is equal to LGA's 'LGA', to form a dataset with 14 variables:
```{r}
airbnb_LGA <- left_join(airbnb, LGA, by = c("neighbourhood_cleansed" = "LGA"))
head(airbnb_LGA)
dim(airbnb_LGA)
```




## Understand 

```{r}
str(airbnb_LGA)
```
\
As seen above, the imported data structure is a data frame, however the majority of the variables are not the correct variable type. Before performing data type conversions, some of the variables will be renamed to make more sense:
```{r}
colnames(airbnb_LGA)[4] <- "LGA"
colnames(airbnb_LGA)[3] <- "location"
head(airbnb_LGA)
```
\
Below, the following conversions take place:

* id:                 Numeric to Factor
* host_since:         Character to Date
* property_type:      Character to Factor
* instant_bookable:   Character to Factor
* LGA:                Character to Factor
* accommodates:       Character to Ordinal Factor
* bathrooms:          Character to Ordinal Factor
* bedrooms:           Character to Ordinal Factor
* guests_included:    Character to Ordinal Factor
* price:              Character to Numeric
* LGA_Area_km2:       Character to Numeric
* LGA_Density:        Character to Numeric

For the last 3 variables, readr's parse_number function is used to remove dollar signs and commas in order to convert the characters into numerics:
```{r}
airbnb_LGA$id <- as.factor(airbnb_LGA$id)
airbnb_LGA$host_since <- ymd(airbnb_LGA$host_since)
airbnb_LGA$property_type <- factor(airbnb_LGA$property_type)
airbnb_LGA$instant_bookable <- factor(airbnb_LGA$instant_bookable)
airbnb_LGA$LGA <- factor(airbnb_LGA$LGA)

unique(airbnb_LGA$accommodates)
airbnb_LGA$accommodates <- factor(airbnb_LGA$accommodates,
                                  levels = c("1","2","3","4","5","6","7","8","9","10","11","12","13","14","15"),
                                  ordered = T)
levels(airbnb_LGA$accommodates)

unique(airbnb_LGA$bathrooms)
airbnb_LGA$bathrooms <- factor(airbnb_LGA$bathrooms,
                                  levels = c("0","0.5","1","1.5","2","2.5","3","3.5","4","4.5","5","5.5",
                                             "6","6.5","7","7.5","8","9","10","14"),
                                  ordered = T)
levels(airbnb_LGA$bathrooms)

unique(airbnb_LGA$bedrooms)
airbnb_LGA$bedrooms <- factor(airbnb_LGA$bedrooms,
                               levels = c("0","1","2","3","4","5","6","7","8","9","10","11"),
                               ordered = T)
levels(airbnb_LGA$bedrooms)

unique(airbnb_LGA$guests_included)
airbnb_LGA$guests_included <- factor(airbnb_LGA$guests_included,
                              levels = c("1","2","3","4","5","6","7","8","9","10","11","12","14","15","16"),
                              ordered = T)
levels(airbnb_LGA$guests_included)

airbnb_LGA$price <- parse_number(airbnb_LGA$price)
airbnb_LGA$LGA_Area_km2 <- parse_number(airbnb_LGA$LGA_Area_km2)
airbnb_LGA$LGA_Density <- parse_number(airbnb_LGA$LGA_Density)
```
\
The following is the dataset after the variable type conversions:
```{r}
str(airbnb_LGA)
```




##	Tidy & Manipulate Data I 

The dataset is in a tidy format since the following apply:

* Each variable has its own column.
* Each observation has its own row.
* Each value has its own cell.




##	Tidy & Manipulate Data II 

Below, two new variables are created via the mutate function (from dplyr). The variable 'price_per_guest' is created by dividing the price ('price') by the number of guests included ('guest_included'). As 'guests_included' was converted to a factor earlier, it is temporarily coerced into a numeric in order for the arithmetic to work. The variable 'LGA_population' is created by multiplying the LGA area (LGA_Area_km2) with the LGA population density (LGA_Density):
```{r}
airbnb_LGA <- airbnb_LGA %>%
  mutate(price_per_guest = round(price/as.numeric(guests_included), 2),
         LGA_population = LGA_Area_km2*LGA_Density)
head(airbnb_LGA)
```


```{r}
airbnb_LGA <- airbnb_LGA %>% separate(location, into = c("Suburb", "State", "Country"), sep = ",")
head(airbnb_LGA)

```


##	Scan I 

Below, the sapply function is used to scan all variables in the dataset for NA values:
```{r}
sapply(airbnb_LGA, function(x) sum(is.na(x)))
sum(is.na(airbnb_LGA))
```
\
As shown above, there are 108 NA values in the entire dataset, which is a very small number compared to the total number of observations, hence these NA observations can be removed:
```{r}
airbnb_LGA <- na.omit(airbnb_LGA)

sapply(airbnb_LGA, function(x) sum(is.na(x)))
```
\
Below is a function to scan all numerical variables for special values such as infinite numbers or NaNs, which is applied to the entire dataset. The results indicate that there are no special values in this dataset:
```{r}
special <- function(x){if (is.numeric(x)) sum(is.infinite(x) | is.nan(x))}
sapply(airbnb_LGA, special)
```
\
Check inconsistency from separated variable, "Suburb", "State", and "Country"

```{r}
unique (airbnb_LGA$Suburb)

```

change all observation to lower case using stringr function

```{r}
airbnb_LGA$Suburb <- str_to_lower(airbnb_LGA$Suburb)
```


```{r}
unique (airbnb_LGA$State)
unique (airbnb_LGA$Country)
```

The actual airbnb location is in Victoria and Australia, so other than that will be changed to "VIC" and "AU"

```{r}
airbnb_LGA$State <- replace(airbnb_LGA$State, airbnb_LGA$State != "VIC", "VIC")

airbnb_LGA$Country <- replace(airbnb_LGA$Country, airbnb_LGA$Country != "AU", "AU")
```

\
Using the editrules package, the following code sets some basic rules to scan for inconsistencies. The rules defined below are to make sure that no numerical value is a negative number. For the price variable however, the rule makes sure that no value is below 5 as for the purpose of this project, any Airbnb accommodation for less than $5 per night is considered a mistake or is not a legitimate listing. There are 22 violations under the rule number 1 (price >= 5), those of which can be seen below:

```{r warning=FALSE, message=FALSE}
library(editrules)
```

```{r}
rules <- editset(c("price >= 5",
                   "price_per_guest >= 0",
                   "number_of_reviews >= 0",
                   "LGA_Area_km2 >= 0",
                   "LGA_Density >= 0",
                   "LGA_population >= 0"))
rules
summary(violatedEdits(rules,airbnb_LGA))
which(violatedEdits(rules,airbnb_LGA))
```

\
The observations with violated rules are removed, resulting in zero violated rules:
```{r}
airbnb_LGA <- airbnb_LGA[-(which(violatedEdits(rules,airbnb_LGA))),]
summary(violatedEdits(rules,airbnb_LGA))
```
\
Another inconsistency within the dataset is that some values of the guests_included variable are larger than the accommodates variable, which is impossible as 'accommodates' is the maximum number of guests allowed to stay in the accommodation. Below, it is shown that 108 observations have guests_included being larger than accommodates.
```{r}
summary(as.numeric(airbnb_LGA$guests_included) > as.numeric(airbnb_LGA$accommodates))
which(as.numeric(airbnb_LGA$guests_included) > as.numeric(airbnb_LGA$accommodates))
```
\
To deal with this, an ifelse statement is used to determine the observations with this inconsistency, which are then imputed with the the value from the accommodates variable:
```{r}
airbnb_LGA <- airbnb_LGA %>%
  mutate(guests_included = ifelse(as.numeric(guests_included) <= as.numeric(accommodates), guests_included, accommodates))

summary(as.numeric(airbnb_LGA$guests_included) > as.numeric(airbnb_LGA$accommodates))
```

##	Scan II

Before scanning for outliers, an upper threshold will be applied to the price variable in order to remove listings with obvious pricing errors. A search on Airbnb listings in Victoria show that the most luxurious mansions list for no more than \$3,500 per night, while a wedding reception accommodation is listed for around \$5,000. Listings higher than \$5,000 per night are usually small dwellings, which are the result of pricing errors or that they're not legitimate listings. The following shows that the dataset contains prices that go up into the \$12,000 range. Any listing with prices above $5,000 per night is removed:
```{r}
max(airbnb_LGA$price)
which(airbnb_LGA$price > 5000)
airbnb_LGA <- airbnb_LGA[-(which(airbnb_LGA$price > 5000)),]
max(airbnb_LGA$price)
```
\
In terms of scanning the numerical variables for outliers, 'LGA_Area_km2', 'LGA_Density', and 'LGA_population' will not be scanned since these values are tied to the factor variable 'LGA' which contain 30 factors, hence only giving 30 unique values for each of these variables. Any outliers resulting from these variables will not be removed or imputed as they simply demonstrate that some LGAs may have significantly lower or higher populations relative to each other. The numerical variables that will be scanned are 'price', 'price_per_guest', and 'number_of_reviews'.

Below, it is shown that all 3 of the variables are right-skewed due to the fact that Airbnb prices and the number of reviews cannot be negative, but can be extremely large. Because of this, Tukey's method (boxplot) will be used to detect outliers as it does not require the assumption of normality. It is shown that the number of outliers for each variable is quite large, with all of them making up more than five percent of the total observations. This means that imputation will be used to deal with these outliers as opposed to removal. It is also important to note that the following outliers probably do not represent errors. Under normal circumstances, these outliers will be left alone as they represent real and correct values that describe the distribution of Airbnb prices and number of reviews, however for the purpose of demonstration, they will be dealt with. 

**Price:**
```{r}
hist(airbnb_LGA$price, breaks = 200, main="Price Per Night", xlab="Price")
price_box <- airbnb_LGA$price %>%
  boxplot(main="Price Per Night", ylab="Price", col = "lightgrey")
summary(airbnb_LGA$price)
length(price_box$out)
length(price_box$out)/nrow(airbnb_LGA)
```
\
**Price Per Guest:**
```{r}
hist(airbnb_LGA$price_per_guest, breaks = 200, main="Price Per Guest", xlab="Price")
price_per_guest_box <- airbnb_LGA$price_per_guest %>%
  boxplot(main="Price Per Guest", ylab="Price", col = "lightgrey")
summary(airbnb_LGA$price_per_guest)
length(price_per_guest_box$out)
length(price_per_guest_box$out)/nrow(airbnb_LGA)
```
\
**Number of Reviews:**
```{r}
hist(airbnb_LGA$number_of_reviews, breaks = 200, main="Number of Reviews", xlab="Number")
number_of_reviews_box <- airbnb_LGA$number_of_reviews %>%
  boxplot(main="Number of Reviews", ylab="Number", col = "lightgrey")
summary(airbnb_LGA$number_of_reviews)
length(number_of_reviews_box$out)
length(number_of_reviews_box$out)/nrow(airbnb_LGA)
```
\
For this dataset, imputation is done via imputing the outliers with the median since the distributions are extremely skewed. Winsorising was considered, however it would've resulted in a large spike in the nearest neighbour of the distribution given the large number of outliers. The median method used also resulted in a large spike in the median of the distribution, so neither solution is ideal (later in the "Transform" section another technique will be explored: the variable 'price_duplicate' is created in the code below for later transformation). For each variable, the following code replaces the outliers found above (that were found using the boxplot function) with the median of its variable:
```{r}
airbnb_LGA <- airbnb_LGA %>%
  mutate(price_duplicate = price)

airbnb_LGA$price[which(airbnb_LGA$price < price_box$stats[1] | airbnb_LGA$price > price_box$stats[5])] <- 
  median(airbnb_LGA$price)
hist(airbnb_LGA$price, breaks = 200, main="Price - Imputed", xlab="Price")
summary(airbnb_LGA$price)

airbnb_LGA$price_per_guest[which(airbnb_LGA$price_per_guest < price_per_guest_box$stats[1] | airbnb_LGA$price_per_guest > price_per_guest_box$stats[5])] <- 
  median(airbnb_LGA$price_per_guest)
hist(airbnb_LGA$price_per_guest, breaks = 200, main="Price Per Guest - Imputed", xlab="Price")
summary(airbnb_LGA$price_per_guest)

airbnb_LGA$number_of_reviews[which(airbnb_LGA$number_of_reviews < number_of_reviews_box$stats[1] | airbnb_LGA$number_of_reviews > number_of_reviews_box$stats[5])] <- 
  median(airbnb_LGA$number_of_reviews)
hist(airbnb_LGA$number_of_reviews, breaks = 200, main="Number of Reviews - Imputed", xlab="Price")
summary(airbnb_LGA$number_of_reviews)
```




##	Transform 

In this section, the 'price_duplicate' variable is transformed and normalised in order to perform a z-score outlier detection. The following uses the 'forecast' package in order to perform a BoxCox transformation. Comparing the tranformed distribution with original distribution, the transformed distribution is much more normal:
```{r}
hist(airbnb_LGA$price_duplicate, breaks = 100, main="Price - Original", xlab="Price")
price_boxcox <- BoxCox(airbnb_LGA$price_duplicate,lambda = "auto")
hist(price_boxcox, breaks = 100, main="Price - Transformed", xlab="Price")
```
\
Below, the tranformed distribution (price_boxcox) can be standardised given its approximately normal distribution. The number of outliers (values that exceed 3 standard deviations) are 88 in this case. Much less than Tukey's method when applied to the non-transformed price variable (which had 1,399 outliers):
```{r}
z_scores <- scale(price_boxcox, center = T, scale = T)
summary(z_scores)
length(which(abs(z_scores) > 3))
```



<br>
<br>
